`make.modifiers`

### 2023-04-27 inst/gitbook/01_getting_started.Rmd
fixed a couple of comments to the manual

@@

First we need to create a `condition` function to trigger this (non) speciation event.
We can do that by specifying our `condition` function (here the `staying.alive` function) to apply our modification.
For that we can use the `parent.traits` utility function that is optimised for accessing traits in the birth-death process (but you can of course write your own).
This function takes the `trait.values` and `parent.lineage` arguments, two arguments that you can leave named as they are to facilitate `treats`' understanding of what you want to assess:

@@

We can then provide these two functions (the condition `negative.ancestor` and how to modify the speciation event when this condition is met `staying.alive`).
If you are an advances `treats` user, you can design your own `speciation` function but if you just want to use a normal `speciation` function, you can use the default one from `treats` called `speciation`.

@@

```{r}
set.seed(1)
## Simulate disparity and diversity
biased_data <- treats(bd.params = extinction_02,
                    stop.rule = living_20,
                    traits    = BM_trait,
                    modifiers = negatives_extinct)
biased_data
```

@@

```{r, fig.height = 6, fig.width = 12}
par(mfrow = c(1,2))
plot(default_data, main = "default results")
plot(biased_data, main = "results with the modifier")
```

>>>>

<!-- ## An illustrative example -->

@@

First we need to create a `condition` function to trigger this (non) speciation event.
We can do that by specifying our `condition` function (here the `going.extinct` function) to apply our modification.
For that we can use the `parent.traits` utility function that is optimised for accessing traits in the birth-death process (but you can of course write your own).
This function takes the `trait.values` and `parent.lineage` arguments, two arguments that you can leave named as they are to facilitate `treats`' understanding of what you want to assess:

@@

We can then provide these two functions (the condition `negative.ancestor` and how to modify the speciation event when this condition is met `going.extinct`).
If you are an advances `treats` user, you can design your own `speciation` function but if you just want to use a normal `speciation` function, you can use the default one from `treats` called `speciation`.

@@

```{r}
set.seed(1)
## Simulate disparity and diversity
modified_data <- treats(bd.params = extinction_02,
                        stop.rule = living_20,
                        traits    = BM_trait,
                        modifiers = negatives_extinct)
modified_data
```

@@

```{r, fig.height = 6, fig.width = 12}
par(mfrow = c(1,2))
plot(default_data, main = "default results")
plot(modified_data, main = "results with the modifier")
```

### 2023-04-27 inst/gitbook/02_make.traits.Rmd
fixed a couple of comments to the manual

In `treats`, traits are simulating by providing a `traits` object to the `traits` argument.
This object is generated using `make.traits`.

> **What is a trait in biology?** This is one of these epistemological questions like what _is_ a species and is not covered in this package.
For the purpose of _this manual_, a **biological trait** can be any coherent character of any number of dimensions which may or may not be independent of other traits.
Basically anything can be a trait if it's defined by the user (e.g. one trait can be a 20 dimensional Brownian process that's correlated to a 1D OU process - which is another trait).
For the purpose of the syntax in _this package_ a **trait in `treats`** is a column in the output trait matrix.
For example, one might be interested in a biological trait that is "beak shape" which is defined as one complex biological trait that is the Procrustes superimposition of 4 3D landmarks and 300 3D semi-landmarks.
This single "beak shape" **biological trait** is then composed of 912 **traits in `treats`** (i.e. a matrix of $(4+300)\times3$ columns).

@@

Note that the number of traits are distributed per processes.
If the traits contains multiple processes, the number of traits are distributed per processes:

>>>>

In `treats`, traits are simulating by providing a `traits` object to the `traits` argument.
This object is generated using `make.traits`.
In `treats` a trait designate any coherent character of any number of dimensions which may or may not be independent of other traits.
For example, it can be a uni-dimensional variable uncorrelated to other traits or a multi-dimensional one with different amounts of correlation.

@@

Note that the number of traits are applied for each process by default.
However, you can apply different number of traits for different processes:

### 2023-04-27 inst/gitbook/03_make.modifiers.Rmd
fixed a couple of comments to the manual

```{r}
## A modifier for removing tips with negative values
bias.trait <- function(trait.values, lineage) {
    if(parent.traits(trait.values, lineage) < 0) {
        ## Go extinct!
        return(FALSE)
    } else {
        ## Speciate!
        return(TRUE)
    }
}

## Creating the modifier
biased_trait <- make.modifiers(speciation = bias.trait)

## Simulate a tree and trait with the modifier
set.seed(1)
biased_trait_treats <- treats(bd.params = bd_params,
                          stop.rule = stop_rule,
                          traits    = BM_trait,
                          modifiers = biased_trait)

## Plotting the differences in trees and traits
par(mfrow = c(1, 2))
plot(default_treats, main = "Default trait and tree")
plot(biased_trait_treats, main = "Biased trait and tree")
```

>>>>

```{r}
## A modifier for removing tips with negative values
modify.trait <- function(trait.values, lineage) {
    if(parent.traits(trait.values, lineage) < 0) {
        ## Go extinct!
        return(FALSE)
    } else {
        ## Speciate!
        return(TRUE)
    }
}

## Creating the modifier
modified_trait <- make.modifiers(speciation = modify.trait)

## Simulate a tree and trait with the modifier
set.seed(1)
modified_trait_treats <- treats(bd.params = bd_params,
                          stop.rule = stop_rule,
                          traits    = BM_trait,
                          modifiers = modified_trait)

## Plotting the differences in trees and traits
par(mfrow = c(1, 2))
plot(default_treats, main = "Default trait and tree")
plot(modified_trait_treats, main = "Biased trait and tree")
```
